---
title: "Quick Start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installation

### Remote installation

When the package is publically available from GitHub *cub* repository:

```{r eval = FALSE}
devtools::install_github(
  'MoBaBioinf/cph-cub', 
  dependencies = TRUE, 
  build_vignettes = TRUE
)
```

### Local installation

- Clone the package to your local computer;
- Open in RStudio;

From R CLI run

```{r eval = FALSE}
devtools::install(dependencies = TRUE, build_vignettes = TRUE)
```




# Input data

All the functions in this package are designed to work with `Biostrings::DNAString` and `Biostrings::DNAStringSet` data types.

# Basic workflow

## Extract codons from CDSs

The package is shipped with the CDSs from *P. aeruginosa*  *PAO1* and *PA14* strains:

```{r message = FALSE}
library(cub)

dna01 <- PAO1CDS
dna02 <- PA14CDS

show(dna01)
```
