---
title: "Quick Start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cub)
```

# Installation

The package is available from GitHub *utubun/cub* repository. To install it you will require `devtools` library to be installed. 

```{r eval = FALSE}
devtools::install_github('utubun/cub', dependencies = TRUE, build_vignettes = TRUE)
```

This makes all the functionality and data sets of the package available to you.

# Input data

All the functions in this package are designed to work with `Biostrings::DNAString` and `Biostrings::DNAStringSet` data types.

# Basic workflow

## Extract codons from CDSs

The package is shipped with the CDSs from *P. aeruginosa*  *PAO1* and *PA14* strains:

```{r message = FALSE}
#library(cub)

pao1 <- PAO1CDS

pao1
```

Next we will convert any gene of interest into `Cub` object, to access the tools for *codon usage bias* analysis.

```{r}
cub01 <- cub('Gene A', dna[[1]])
```

